---
title: "pax"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE)
```

### Premable

... pax: population analysis in unix. This is a resuscitation of an old name - but likely going to do more.

Have you wished for a painless way to generate survey indices. Give the following a try.

### Needed packages

```{r, eval = FALSE}
devtools::install_github("fishvice/husky", dependencies = FALSE) # A temporary measure
devtools::install_github("fishvice/pax", dependencies = FALSE)
```

In addition it is assumed you have the __fjolst__-package.

### Loading needed libraries

```{r}
library(tidyverse)
library(fjolst)
library(pax)
```

### An example

#### Specify the species

```{r}
SPECIES <- 1
KYN <- NULL
```

#### SMB all stations

OK, here you have to have some bacground knowledge. But we select "all" SMB stations:

```{r}
Station <- 
  husky::STODVAR %>% 
  bind_rows() %>% 
  filter(tognumer %in% 1:39) %>% 
  select(id = synis.id, year = ar, towlength = toglengd, strata = newstrata) %>% 
  mutate(towlength = pax:::trim_towlength(towlength),
         mult = 1) %>% 
  arrange(id)
```

And get the appropriate strata as well (the old strata, which is not the same as the very old strata):

```{r}
Stratas <-
  husky::stratas_df %>% 
  select(strata, area = rall.area)
```

And then we leave the rest to the `smx:calc_length_indices`-function:

```{r}
res <- calc_length_indices(Station, Stratas, SPECIES)
```

#### Visualization

Extracting the tidy information and preparing the plot:
```{r}
tidy_fixed <-
  res$aggr %>% 
  filter(length == 5) %>% 
  mutate(source = "tidy")
p <- 
  tidy_fixed %>% 
  select(year, cb, cb.cv) %>% 
  ggplot(aes(year, cb)) +
  geom_pointrange(aes(ymin = cb * (1 - cb.cv),
                      ymax = cb * (1 + cb.cv)),
                  colour = "red", lwd = 1) +
  geom_line(colour = "red", lwd = 1) +
  scale_colour_brewer(palette = "Set1") +
  labs(x = NULL, y = NULL,
       title = "Haddock: Spring survey biomass index",
       subtitle = "All stations")
```

Lets add the official calculations for comparison:
```{r, plot1}
attach("/net/hafkaldi/export/u2/reikn/Splus5/SMB/Allindices.RData")
mri <- 
  All.indices %>% 
  filter(species == SPECIES,
         svaedi == "Heild",
         lengd == 5,
         type == "all") %>% 
  select(year = ar, cb = bio.staerri, cb.cv = cv.bio.staerri, type) %>% 
  mutate(source = "mri")

p +
  geom_pointrange(data = mri, 
                  aes(ymin = cb * (1 - cb.cv),
                      ymax = cb * (1 + cb.cv)),
                  colour = "yellow") +
  geom_line(data = mri, colour = "yellow") +
  labs(x = NULL, y = NULL,
       title = "Biomass indices",
       subtitle = "Comparison of the tidyverse (red) and the Bible (grey)")
```

#### Something to worry about?

```{r}
tidy_fixed %>% 
  select(year, cb) %>% 
  left_join(mri %>% select(year, cb2 = cb)) %>% 
  mutate(diff = (cb - cb2)/cb2 * 100) %>% 
  summary()
```

So the tidyverse is 0.2% higher than the mri - some may want to dig into that.

